Response to referee

### 2024-04-15 paper/resp_referee.qmd
small updates

1. Page 3, line 46: you say that it is "equivalently defined" and then provide the conjugate,
writing $e^{i\omega k}$ instead of $e^{-i\omega k}$. "Equivalence" implies something akin to equality, so I would soften this to "alternatively", since the two expressions are not equal.

2. Page 5, line 53: replace "have" with "design", I think, because you are constructing these
filters, right?

3. Section 1.2.1: these results are well-known; the derivatives of the frequency response function
at $\omega = 0$ yield the conditions. So the results could be stated with references, and the
derivations omitted, and this subsection compactified.

4. Page 7, line 40: why say the series is extended by one year? The extensions are for as many
years as needed by the particular filters (typically, half their length in years).

5. Page 8, line 10: omit the $\forall q$ since you have stated this qualification in the previous line.

6. Table 1: I repeat the referees here. The table needs more explication in the text. What is $\omega_1$? Define $\nabla$, Guggemos, etc.

7. Page 13, equation (4): it is hard to understand what $\theta_f$ and $\theta_p$ are, and what 
$\gamma$ and $\delta$ are.
Add more explanation to the text.

8. Page 15, line 36: should be "same length as $v$"

9. Page 15, line 40: I couldn't find the matrix $T$ defined. How is it connected to timeliness?
That is, when you display the entries of the matrix T, you need to make a connection between
those values and the concept of timeliness.

10. Page 15, line 53: I don't see why $v$ in this equation should be defined through the case bracket;
everything starting with "with" could be moved to the following text.

11. Page 17, lines 53-55: This discussion needs more clarity; it seems a bit terse.

12. Appendix B: rather than typing out the R code, it would be better to have the files available
as supplementary material. (Or possibly put in a github repo and provide the links in the
text.)

@@

The authors should provide a clear definition of trend-cycle, $TC_t$,
and irregular, $I_t$. In other words, how is the decomposition
$y_t =TC_t+I_t$ identified? In section 1 they seem to hint at the interpretation
of $TC_t$ as the low-pass component of yt with cutoff frequency
$\pi/6$ (and correspondingly they interpret $I_t$ as the high-pass component).
This point should be clarified, I think.
Related questions are: are $y_t$ and its components stochastic or deterministic?
The treatment in section 1.2.1 suggests that $TC_t$ is
an unknown possibly nonlinear deterministic function of time.
The variance inflation factor $\sum_k \theta_k^2$ implies that $I_t$ is white noise.

2. The idea of minimizing the noise component ($M_\theta(I_t)\simeq 0$) is not
very formal. Do you mean that the unconditional variance is minimized?
Or the long run variance?

3. A definition of turning point would also be needed. Are these ’regular’
points at which the slope of the trend changes sign? Or changepoints?

4. The notation should be unified: why not writing $TC_t = \mu(t)$ and
in section 2 using $t$ instead of $x$? The observation index i hould
be replaced by $t$, as in section 2.1. Similarly, why not using ϵt
directly and start section 1.2 by stating that you assume that yt is
generated as $y_t = \mu(t) + \varepsilon_t$, etc.
Secondly, you should clarify the meaning of (3), where you set
$\mu(t) = u'_t\gamma + z'_t\delta$ (also you assume that the noise component has possibly a non-scalar covariance matrix, do you need the matrix
$D$ for later treatment?
5. You should state that you assume a fixed bandwidth. A nearest
neighbour bandwidth should be interesting to consider: as such
it would contribute more to the reduction of the variance at the
expenses of the phase shift.
6. Is the phase shift related to the notion of bias in local polynomial
smoothing?

@@

- Section 1.1. It would perhaps be useful to provide the expression
of the phase angle, $\varphi(\omega)$ in terms of the real and imaginary part of
the transfer function of the linear filter.
- Section 1.1. The symbol p for the period is already in use from the
previous section, where it denotes the bandwidth and number of
past observations used by the filter.
- Page 4. The definition of the high-pass component (oscillations
in the frequency range from $\pi/6$ to $\pi$ as undesirable oscillations is
controversial.
- Figure 1. In the caption, the ration $I/C$ was never defined before.
The description of the three panels should be better explained in
the main text.
- Section 1.2 makes a reference to X-11 and X-13-ARIMA. The author
assumes that the reader is knowledgeable about seasonal adjustment
methods. Perhaps the authors should state that their
intent is to separate the trend-cycle component from the ’noise’
component.
- Page 7, line 38. ’which extends’ is repeated twice.
- Page 8. What is the meaning of $y_1^*n\dots, y_h^*$ being the implicit forecasts
induced by $w_0$, etc.? The forecasts are obtained from an assumed
polynomial trend model?
- Table 1. Some criteria are labelled as Guggemos, not referenced in
the main text.
- Figures 6-7. The labels of the time axes and the months are in
French.

@@

Those sections are now numbered

@@

Removed

>>>>

> 1. Page 3, line 46: you say that it is "equivalently defined" and then provide the conjugate, writing $e^{i\omega k}$ instead of $e^{-i\omega k}$.  
> "Equivalence" implies something akin to equality, so I would soften this to "alternatively", since the two expressions are not equal.

*It was corrected.*

> 2. Page 5, line 53: replace "have" with "design", I think, because you are constructing these filters, right?

*Indeed, it was corrected.*

> 3. Section 1.2.1: these results are well-known; the derivatives of the frequency response function 
> at $\omega = 0$ yield the conditions. So the results could be stated with references, and the 
> derivations omitted, and this subsection compactified.

> 4. Page 7, line 40: why say the series is extended by one year? The extensions are for as many 
> years as needed by the particular filters (typically, half their length in years).

> 5. Page 8, line 10: omit the $\forall q$ since you have stated this qualification in the previous line.

> 6. Table 1: I repeat the referees here. The table needs more explication in the text. What is $\omega_1$? Define $\nabla$, Guggemos, etc.

*The section 1.3.2 was removed: it's not central to the paper, and it would have made the reading more cumbersome to go into the details of the notations. Table 2 has also been removed *

> 7. Page 13, equation (4): it is hard to understand what $\theta_f$ and $\theta_p$ are, and what $\gamma$ and $\delta$ are. 
> Add more explanation to the text.

> 8. Page 15, line 36: should be "same length as $v$"

> 9. Page 15, line 40: I couldn't find the matrix $T$ defined. How is it connected to timeliness? 
> That is, when you display the entries of the matrix T, you need to make a connection between 
> those values and the concept of timeliness.

> 10. Page 15, line 53: I don't see why $v$ in this equation should be defined through the case bracket; 
> everything starting with "with" could be moved to the following text.

> 11. Page 17, lines 53-55: This discussion needs more clarity; it seems a bit terse.

> 12. Appendix B: rather than typing out the R code, it would be better to have the files available 
> as supplementary material. (Or possibly put in a github repo and provide the links in the 
> text.)

@@

> 1. The authors should provide a clear definition of trend-cycle, $TC_t$, 
> and irregular, $I_t$. In other words, how is the decomposition 
> $y_t =TC_t+I_t$ identified? In section 1 they seem to hint at the interpretation 
> of $TC_t$ as the low-pass component of yt with cutoff frequency 
> $\pi/6$ (and correspondingly they interpret $I_t$ as the high-pass component). 
> This point should be clarified, I think. 
> Related questions are: are $y_t$ and its components stochastic or deterministic? 
> The treatment in section 1.2.1 suggests that $TC_t$ is 
> an unknown possibly nonlinear deterministic function of time. 
> The variance inflation factor $\sum_k \theta_k^2$ implies that $I_t$ is white noise.

> 2. The idea of minimizing the noise component ($M_\theta(I_t)\simeq 0$) is not
> very formal. Do you mean that the unconditional variance is minimized?
> Or the long run variance?

> 3. A definition of turning point would also be needed. Are these ’regular’
points at which the slope of the trend changes sign? Or changepoints?

> 4. The notation should be unified: why not writing $TC_t = \mu(t)$ and
> in section 2 using $t$ instead of $x$? The observation index i hould
> be replaced by $t$, as in section 2.1. Similarly, why not using ϵt
> directly and start section 1.2 by stating that you assume that yt is
> generated as $y_t = \mu(t) + \varepsilon_t$, etc.
> Secondly, you should clarify the meaning of (3), where you set
> $\mu(t) = u'_t\gamma + z'_t\delta$ (also you assume that the noise component has possibly a non-scalar covariance matrix, do you need the matrix $D$ for later treatment?

> 5. You should state that you assume a fixed bandwidth. 
A nearest neighbour bandwidth should be interesting to consider: as such it would contribute more to the reduction of the variance at the expenses of the phase shift.

> 6. Is the phase shift related to the notion of bias in local polynomial
smoothing?

@@

> - Section 1.1. It would perhaps be useful to provide the expression
of the phase angle, $\varphi(\omega)$ in terms of the real and imaginary part of
the transfer function of the linear filter.

> - Section 1.1. The symbol $p$ for the period is already in use from the
previous section, where it denotes the bandwidth and number of
past observations used by the filter.

> - Page 4. The definition of the high-pass component (oscillations
in the frequency range from $\pi/6$ to $\pi$ as undesirable oscillations is
controversial.

> - Figure 1. In the caption, the ration $I/C$ was never defined before.
The description of the three panels should be better explained in
the main text.

> - Section 1.2 makes a reference to X-11 and X-13-ARIMA. The author
assumes that the reader is knowledgeable about seasonal adjustment
methods. Perhaps the authors should state that their
intent is to separate the trend-cycle component from the ’noise’
component.

> - Page 7, line 38. ’which extends’ is repeated twice.

> - Page 8. What is the meaning of $y_1^*n\dots, y_h^*$ being the implicit forecasts induced by $w_0$, etc.? 
The forecasts are obtained from an assumed polynomial trend model?

> - Table 1. Some criteria are labelled as Guggemos, not referenced in the main text.

*The section 1.3.2 was removed: it's not central to the paper, and it would have made the reading more cumbersome to go into the details of the notations. Table 2 has also been removed *

> - Figures 6-7. The labels of the time axes and the months are in
French.

@@

*Those sections are now numbered*

@@

*The quote was removed*
